---
title: 'MakeoverMonday 2019/W16'
output:
  html_document:
    css: style.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(glmnet)



# Import contents ----

contents_orig <- readxl::read_excel("data/Info We Trust _ #MakeoverMonday.xlsx", sheet = "contents")
sections_vec <- unique(contents_orig$section)

# Calculate a number of pages in each section of the book.
contents_orig %>% 
  group_by(section) %>% 
  mutate(n_pages = max(page) - min(page)) %>% 
  distinct(section, n_pages)



# Import words ----

words_orig <- readxl::read_excel("data/Info We Trust _ #MakeoverMonday.xlsx", sheet = "data")
words_orig



# Word transformations ----

# Lemmatization.
words <- words_orig %>% 
  left_join(contents_orig, by = "page") %>% 
  mutate(word_lemma = textstem::lemmatize_words(wordForm),
         section = factor(section, 
                          levels = sections_vec, 
                          labels = paste(1:length(sections_vec), sections_vec, sep = ". ")))

# Thanks to Simon Jackson @drsimonj for sharing the code for ordering categories within facets.
# https://drsimonj.svbtle.com/ordering-categories-within-ggplot2-facets


# View lemmatized words.
words %>% 
  filter(wordForm != word_lemma) %>% 
  count(wordForm, word_lemma, sort = TRUE)

# Transform "numb" back to "number" (side effect of lemmatization).
words <- words %>% 
  mutate(word_lemma = case_when(word_lemma == "numb" ~ "number", 
                                word_lemma == "datum" ~ "data", 
                                TRUE ~ word_lemma))

# Calculate TF-IDF (approx. as number of words in each section is unknown).
sections_per_word <- words %>% 
  group_by(word_lemma) %>% 
  summarize(n_sections = n_distinct(section))

words_in_sections <- words %>% 
  group_by(section) %>% 
  summarize(words_in_section = n())

word_by_section <- words %>% 
  count(section, word_lemma) %>% 
  arrange(section, desc(n)) %>% 
  left_join(sections_per_word, by = "word_lemma") %>% 
  left_join(words_in_sections, by = "section") %>% 
  mutate(idf = log(length(sections_vec) / n_sections, base = exp(1)),
         tf_idf = n / words_in_section * idf)


# Plot word importance as characterized by TF-IDF.

top_10_tfidf <- word_by_section %>% 
  group_by(section) %>% 
  arrange(desc(tf_idf)) %>% 
  slice(1:10) %>% 
  ungroup() %>% 
  # Arrange by facet and word frequency.
  arrange(section, tf_idf) %>% 
  # Add order column of row numbers
  mutate(order = row_number())

subtitle_text <- readLines("subtitle_text.txt")

p_tfidf <- top_10_tfidf %>% 
  ggplot(aes(x = order, y = tf_idf)) +
  geom_col(fill = "#F7C03E") +
  facet_wrap(~section, scales = "free") +
  scale_x_continuous(breaks = top_10_tfidf$order,
                     labels = top_10_tfidf$word_lemma) +
  coord_flip() +
  theme_minimal() +
  labs(title = "Importance of words in RJ Andrew's book 'Info We Trust'",
       subtitle = paste0("\n", str_wrap(subtitle_text, width = 150), "\n", collapse = ""),
       caption = "Makeover in R by Dmitrijs Kass @dmitrijsk | Data source: https://data.world/makeovermonday/2019w16",
       x = NULL, 
       y = NULL) +
  theme(strip.background = element_blank(), 
        panel.grid = element_blank(), 
        plot.background = element_rect(fill = "#2C5696"), 
        strip.text = element_text(colour = "white", face = "bold", size = 12), 
        axis.text = element_text(colour = "white"), 
        title = element_text(colour = "white", size = 15), 
        plot.subtitle = element_text(colour = "white", size = 9.5),
        axis.text.x = element_blank(),
        axis.text.y = element_text(hjust = 0, size = 11), 
        plot.caption = element_text(size = 8, hjust = 0),
        panel.spacing.x = unit(2,"line"), 
        panel.spacing.y = unit(1.5,"line"))

ggsave(filename = "images-raw/tf_idf.png", height = 7, width = 10)
```

This week focuses on the words in RJ Andrew's book ["Info We Trust"](https://infowetrust.com/). The original viz is a [word cloud](https://twitter.com/TriMyData/status/1117386318391197698). Data is available [here](https://data.world/makeovermonday/2019w16). My makeover focuses on the importance of words in six sections of the book. Word importance may be useful in highlighting differences in the content of the book's sections. Data analysis and visualization was implemented in R. Code with explanations will be added later.

<br>

```{r echo = FALSE, fig.align='left', fig.height=7, fig.width=10}
p_tfidf
```
